Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-02-03 Grassmannian_cells.tex
Section 3.2 edits (done?)

@@

The next step is to introduce notation in order to locate truncated preprojectives as quotients of other truncated preprojectives in the universal covering.
\begin{definition}
  For $I\subsetneq\{1,\ldots,n\}$, write $I^c=\{1,\ldots,n\}\setminus I$ for the complementary subset.
  A sequence of subsets $\bfI=(I_0,I_1,\ldots,I_k)$, $k\ge1$, in $\{1,\ldots,n\}$ is \emph{admissible} if the following hold:
  \begin{enumerate}
    \item $|I_0|=n-1$ and $|I_l|=n-2$ for $1\le l\le k-1$;
    \item the sets $I'_l$ defined recursively by $I'_0=I_0$ and $I'_{l+1}:=(I'_l)^c\cup I_{l+1}$ for $0\le l\le k$ satisfy $(I'_l)^c\cap I_{l+1}=\varnothing$ for $0\le l\le k$.
  \end{enumerate}
  Here we take $I_{k+1}=\varnothing$.

@@

By Lemma~\ref{le:properties}.(1), we can quotient out the lifted preprojectives successively, i.e.\ for any proper subsets $J\subsetneq I\subsetneq\{1,\ldots,n\}$ we have a short exact sequence
\[\sesm{\bigoplus_{i\in I\setminus J}\tilde P_{m,i}}{\tilde P_{m+1}^J}{\tilde P_{m+1}^I}{\pi_{m+1}^{J,I}}.\]
For any sequence of proper subsets $K\subsetneq J\subsetneq I\subsetneq\{1,\ldots,n\}$, the quotient maps satisfy $\pi_{m+1}^{K,I}=\pi_{m+1}^{J,I}\circ\pi_{m+1}^{K,J}$ and $\pi_{m+1}^I=\pi_{m+1}^{J,I}\circ\pi_{m+1}^J$.
For $w\in W_n$, we write $\pi_{m+1,w}^{I,J}:\tilde P_{m+1,w}^I\to\tilde P_{m+1,w}^J$ for the translated morphism with similar notation for truncated preprojectives $\tilde P_{m+1-k,\ui}^I$ for $\ui\in A_1^{(k)}$ and $I\subsetneq\{1,\ldots,n\}$.
\begin{lemma}
  \label{le:truncated quotients}
  For $m\ge3$, the following hold:
  \begin{enumerate}
    \item For $I\subsetneq\{1,\ldots,n\}$ with $|I|=n-1$ and $I^c=\{j\}$, we have an isomorphism 
      \[\tilde P_{m+1}^I\cong\tilde P_{m,j}^{I^c}=\tilde P_{m,j}/\tilde P_{m-1,(j,j)}.\]
    \item Consider an admissible sequence $\bfI=(I_0,\ldots,I_k)$ with $k<m$ and write $(I'_l)^c=\{i_l\}=:J_l$ for $0\le l\le k-1$.
      For $\ui=(i_0,\ldots,i_{k-1})$, there exists a commutative diagram:
      \begin{equation}
        \label{dia:truncated coherence}
        \xymatrix@R40pt@C70pt{\tilde P_{m+1} \ar^(.4){\pi_{m+1}^{I_0}}[d] & \ar[l] \tilde P_{m,i_0} \ar_(.4){\pi_{m,i_0}^{J_0}}[dl]\ar^(.4){\pi_{m,i_0}^{I'_1}}[d] & \ar[l] \tilde P_{m-1,(i_0,i_1)} \ar_(.4){\pi_{m-1,(i_0,i_1)}^{J_1}\;}[dl]\ar^(.4){\pi_{m-1,(i_0,i_1)}^{I'_2}}[d] & \ar[l] \ar[dl] \cdots & \ar[l] \tilde P_{m+1-k,\ui} \ar_(.4){\pi_{m+1-k,\ui}^{J_{k-1}}\;}[dl]\ar^(.4){\pi_{m+1-k}^{I'_k}}[d]\\
        \tilde P_{m+1}^{I'_0}\ar^{\pi_{m,i_0}^{J_0,I'_1}}[r] & \tilde P_{m,i_0}^{I'_1}\ar^{\pi_{m-1,(i_0,i_1)}^{J_1,I'_2}}[r] & \tilde P_{m-1,(i_0,i_1)}^{I'_2}\ar^(.6){\pi_{m-2,(i_0,i_1,i_2)}^{J_2,I'_3}}[r] & \cdots\ar^{\pi_{m+1-k,\ui}^{J_{k-1},I'_k}}[r] & \tilde P_{m+1-k,\ui}^{I'_k}}
      \end{equation}
      where the composed map $\pi_{m+1}^\bfI:\tilde P_{m+1}\to\tilde P_{m+1-k,\ui}^{I'_k}$ is unique up to scaling and therefore we write
      \[\tilde P_{m+1}^\bfI:=\tilde P_{m+1-k,\ui}^{I'_k}.\]
      These truncated preprojective representations satisfy 
      \[\tilde P_{m+1}^\bfI=\tilde P_{m,i_0}^{\delta\bfI}=\cdots=\tilde P_{m+1-k,\ui}^{\delta^k\bfI}.\]
  \end{enumerate}
\end{lemma}
\begin{proof}

For part (1), we consider the following commutative diagram, where the square on the left hand side is the corresponding Auslander-Reiten sequence:
\[\xymatrix{0\ar[r]&\tau\tilde P_{m+1}=\tilde P_{m-1} \ar[r]\ar[d] & \tilde P_{m,j}\ar[d]\ar[r]& \tilde P_{m+1}^I\ar[r]\ar@{=}[d]& 0\\
  0\ar[r] &\bigoplus_{i\in I}\tilde P_{m,i}\ar[r] &  \tilde P_{m+1}\ar[r] & \tilde P_{m+1}^I\ar[r] & 0}\]

%Thus we see that $\tilde P_{m}^{\ui}\cong \tilde P_{m-1}^{(l)}$ for some $l\in\{1,\ldots,n\}$. But now it is straightforward to check that $l=n$. \footnote{more precise?} 
For $m=3$, it is straightforward that the injection of the lower row factors over $\tilde P_{1,(j^c,j^c)}$ which means that we get the same commutative diagram when replacing $\tilde P_{1}$ by $\tilde P_{1,(j^c,j^c)}$. Thus we have $\tilde P_{3}^{\uj}\cong \tilde P_{2,j^c}^{\uj^c}/\tilde P_{1,(j^c,j^c)}=\tilde P_{2,j^c}^{(j^c)}$. Applying the identities \eqref{eq:recursive covers} the claim follows for $m\geq 3$.

The second statement is now obtained recursively when initially considering the commutative diagram from above replacing $\uj$ by $\underline{j_1}$. We have $\delta^{k-1}{\bf j}(k)=(j_{k-1}^c,j_{k,1},\ldots,j_{k,n-1})$ and $\delta^{k}{\bf j}(k)=(j_k^c)$. We successively obtain commutative diagrams
\[\xymatrix{0\ar[r] &\bigoplus_{i=1}^{n-2}\tilde P_{m-k,(j_1^c,\ldots,j_{k-1}^c,j_{k,i})}\ar[r] &  \tilde P^{(j_{k-1}^c)}_{m-k+1,(j_1^c,\ldots,j_{k-1}^c)}\ar[r] & \tilde P_{m-k+1,(j_1^c,\ldots,j_{k-1}^c)}^{\delta^{k-1}{\bf j}(k)}\ar[r] & 0 \\
  0\ar[r]&\tilde P_{m-(k+1),(j_1^c,\ldots,j_{k}^c,j_k^c)} \ar[r]\ar[u] & \tilde P_{m-k,(j_1^c,\ldots,j_{k}^c)}\ar[u]\ar[r]& \tilde P_{m-k,(j_1^c,\ldots,j_{k}^c)}^{(j_{k}^c)}\ar[r]\ar@{=}[u]& 0}\]

Finally, we consider the short exact sequence:

\[\xymatrix{0\ar[r] &\bigoplus_{i=1}^{s}\tilde P_{m-r,(j_1^c,\ldots,j_{r-1}^c,j_{r,i})}\ar[r] &  \tilde P^{(j_{r-1}^c)}_{m-r+1,(j_1^c,\ldots,j_{r-1}^c)}\ar[r] & \tilde P_{m-r+1,(j_1^c,\ldots,j_{r-1}^c)}^{{\bf j}}\ar[r] & 0 }.\]
Using Lemma \ref{le:properties}.(2), we recursively obtain	$\Hom(\tilde P_m,\tilde P^{\mathbf j}_{m-r+1})\cong\Hom(\tilde P_m,\tilde P_{m-k,(j_1^c,\ldots,j_{k}^c)}^{(j_{k}^c)})=\CC$.
%As $\Hom(\tilde P_{m-k,(j_1^c,\ldots,j_{k-1}^c,j_{k,i})},\tilde P_{m-k+1,(j_1^c,\ldots,j_{k-1}^c)})=\kk$ for all $k=2,\ldots,r$ and $i=1,\ldots,n-1$, applying the second statement, we obtain that the quotient $\tilde P_{m-r+1}^{\bf j}$ is uniquely determined.
	\end{proof}

The lemma implies that there is a morphism by $\pi_m^{\bf j}:\tilde P_m\to \tilde P_{m-r+1}^{\bf j}$ which is unique up to scalars. Let $\underline{j_1}=(j_{1,1},\ldots,j_{1,n-1})$ and $\uj=(j_{1,1},\ldots,j_{1,l})\in I_l$. It is important for us that this morphism factors through the truncated preprojective $\tilde P_m^{\uj}$. We denote the quotient morphism from $\tilde P_m^{\uj}$ to $\tilde P_{m-r+1}^{\bf j}$ by $\pi_{m,\uj}^{\bf j}$. We want to apply $\pi_m^{\bf j}$ also to translated representations $w.\tilde P_m$, mostly for $w\in\{a_i,a_i^{-1}\}$. Therefore, we need to apply a translated version of this morphism. In abuse of notation, we also denote these morphisms by $\pi_m^{\bf j}$ and $\pi_{m,\uj}^{\bf j}$.

Our construction immediately yields the following lemma.
\begin{lemma}\label{projections}
Let ${\bf j}=(\underline{j_1},\ldots,\underline{j_r})$ with $r\leq m-1$ and $\underline{j_r}\in I_s$ be an admissible sequence with entries $j_{k,i}$ for $k=1,\ldots r$. Then we have
\[\pi_m^{\bf j}=\pi_{m-r+1,(j_{r-1}^c)}^{(j_{r-1}^c,\underline{j_r})}\circ\ldots\circ\pi^{(j_{2}^c,\underline{j_2})}_{m-1,(j_1^c)}\circ\pi_m^{\underline{j_1}}.\]
Moreover, for every sequence $(l_1,\ldots,l_{n-2})\in I_{n-2}$, we have
\[\pi_{m-k+1,(j_{k-1}^c)}^{(j_{k-1}^c,\underline{j_k})}=\prod_{i=1}^{n-2} \pi^{(j_{k-1}^c,j_{k,l_1},\ldots, j_{k,l_{i}})}_{m-k+1,(j_{k-1}^c,j_{k,l_1},\ldots,j_{k,l_{i-1}})}\]
for $k=2,\ldots,r-1$. The analogous statements hold for $\pi_m^{\underline{j_1}}$ and  $\pi_{m-r+1,(j_{r-1}^c)}^{(j_{r-1}^c,\underline{j_r})}$ when replacing $n-2$ by $n-1$ and $s-1$ respectively.
\end{lemma}

@@

Thus it remains to show that this cell decomposition of $\Gr^{\tilde Q}_\tbfe(\tilde P_{m}^{\uj'})$ is compatible with $\pi_{m,\uj'}^{\bf j}$ where ${\bf j}=(\underline{j_1},\underline{j_2},\ldots,\underline{j_r})$ with $\underline{j_1}=(\uj',j_{1,l},\ldots,j_{1,n-1})\in I_{n-1}$ and $j_{1,l}=j_l$. Recall that we have $\pi_{m,\uj'}^{\bf j}=\pi_{m,\uj}^{\bf j}\circ\pi_{m,\uj'}^{j_l}$ by Lemma \ref{projections}.

>>>>

\begin{corollary}
  \label{cor:AR lift}
  For $m\ge2$, the Auslander-Reiten sequence~\eqref{eq:AR sequence} on $K(n)$ lifts to the Auslander-Reiten sequence
  \begin{equation}
    \label{eq:lifted AR sequence}
    \xymatrix{0 \ar[r] & \tilde P_{m-1} \ar^(.4){\tilde\iota_{m-1}}[r] & \bigoplus_{j=1}^n \tilde P_{m,j} \ar[r] & \tilde P_{m+1} \ar[r] & 0}.
  \end{equation}
\end{corollary}

@@

The next step is to introduce notation in order to locate truncated preprojectives as quotients of other truncated preprojectives in the universal covering.
\begin{definition}
  For $I\subsetneq\{1,\ldots,n\}$, write $I^c=\{1,\ldots,n\}\setminus I$ for the complementary subset.
  A sequence of subsets $\bfI=(I_0,I_1,\ldots,I_k)$, $k\ge1$, in $\{1,\ldots,n\}$ is \emph{admissible} if the following hold:
  \begin{enumerate}
    \item $|I_0|=n-1$ and $|I_l|=n-2$ for $1\le l\le k-1$;
    \item the sets $I'_l$ defined recursively by $I'_0=I_0$ and $I'_{l+1}:=(I'_l)^c\cup I_{l+1}$ for $0\le l\le k$ satisfy $(I'_l)^c\cap I_{l+1}=\varnothing$ for $0\le l\le k$.
  \end{enumerate}
  Here we take $I_{k+1}=\varnothing$ so that there is no condition imposed on the set $I_k$.

@@

By Lemma~\ref{le:properties}.(1), we can quotient out the lifted preprojectives successively, i.e.\ for any proper subsets $J\subsetneq I\subsetneq\{1,\ldots,n\}$ we have a short exact sequence
\[\sesm{\bigoplus_{i\in I\setminus J}\tilde P_{m,i}}{\tilde P_{m+1}^J}{\tilde P_{m+1}^I}{\pi_{m+1}^{J,I}}.\]
For any sequence of proper subsets $K\subsetneq J\subsetneq I\subsetneq\{1,\ldots,n\}$, the quotient maps satisfy $\pi_{m+1}^{K,I}=\pi_{m+1}^{J,I}\circ\pi_{m+1}^{K,J}$ and $\pi_{m+1}^I=\pi_{m+1}^{J,I}\circ\pi_{m+1}^J$.
For $w\in W_n$, we write $\pi_{m+1,w}^{I,J}:\tilde P_{m+1,w}^I\to\tilde P_{m+1,w}^J$ for the translated morphism with similar notation for truncated preprojectives $\tilde P_{m+1-k,\ui}^I$ for $\ui\in A_1^{(k)}$ and $I\subsetneq\{1,\ldots,n\}$.
\begin{lemma}
  \label{le:truncated quotients}
  For $m\ge3$, the following hold:
  \begin{enumerate}
    \item For $I\subsetneq\{1,\ldots,n\}$ with $|I|=n-1$ and $I^c=\{j\}$, we have an isomorphism 
      \[\tilde P_{m+1}^I\cong\tilde P_{m,j}^{I^c}=\tilde P_{m,j}/\tilde P_{m-1,(j,j)}.\]
    \item Consider an admissible sequence $\bfI=(I_0,\ldots,I_k)$ with $k<m$ and write $(I'_l)^c=\{i_l\}=:J_l$ for $0\le l\le k-1$.
      For $\ui=(i_0,\ldots,i_{k-1})$, there exists a commutative diagram:
      \begin{equation}
        \label{dia:truncated coherence}
        \xymatrix@R40pt@C70pt{\tilde P_{m+1} \ar^(.4){\pi_{m+1}^{I_0}}[d] & \ar[l] \tilde P_{m,i_0} \ar_(.4){\pi_{m,i_0}^{J_0}}[dl]\ar^(.4){\pi_{m,i_0}^{I'_1}}[d] & \ar[l] \tilde P_{m-1,(i_0,i_1)} \ar_(.4){\pi_{m-1,(i_0,i_1)}^{J_1}\;}[dl]\ar^(.4){\pi_{m-1,(i_0,i_1)}^{I'_2}}[d] & \ar[l] \ar[dl] \cdots & \ar[l] \tilde P_{m+1-k,\ui} \ar_(.4){\pi_{m+1-k,\ui}^{J_{k-1}}\;}[dl]\ar^(.4){\pi_{m+1-k}^{I'_k}}[d]\\
        \tilde P_{m+1}^{I'_0}\ar^{\pi_{m,i_0}^{J_0,I'_1}}[r] & \tilde P_{m,i_0}^{I'_1}\ar^{\pi_{m-1,(i_0,i_1)}^{J_1,I'_2}}[r] & \tilde P_{m-1,(i_0,i_1)}^{I'_2}\ar^(.6){\pi_{m-2,(i_0,i_1,i_2)}^{J_2,I'_3}}[r] & \cdots\ar^{\pi_{m+1-k,\ui}^{J_{k-1},I'_k}}[r] & \tilde P_{m+1-k,\ui}^{I'_k}}
      \end{equation}
      where the composed map $\pi_{m+1}^\bfI:\tilde P_{m+1}\to\tilde P_{m+1-k,\ui}^{I'_k}$ is unique up to scaling and therefore we write
      \[\tilde P_{m+1}^\bfI:=\tilde P_{m+1-k,\ui}^{I'_k}.\]
      These truncated preprojective representations satisfy 
      \[\tilde P_{m+1}^\bfI=\tilde P_{m,i_0}^{\delta\bfI}=\cdots=\tilde P_{m+1-k,\ui}^{\delta^k\bfI}.\]
  \end{enumerate}
\end{lemma}
\begin{proof}
  The Auslander-Reiten sequence~\eqref{eq:lifted AR sequence} gives rise to the following commutative diagram:
  \[\xymatrix{& & 0\ar[d]& 0\ar[d]&\\
    & & \bigoplus_{i\in I} \tilde P_{m,i}\ar[d]\ar@{=}[r]& \bigoplus_{i\in I} \tilde P_{m,i}\ar[d]&\\
    0\ar[r]&\tilde P_{m-1} \ar[r]\ar@{=}[d] & \bigoplus_{i=1}^n \tilde P_{m,i}\ar[d]\ar[r]\ar@{}[dr]|(.7){\lrcorner} & \tilde P_{m+1}\ar[r]\ar[d]& 0\\
    0\ar[r] &\tilde P_{m-1}\ar[r] &  \tilde P_{m,j}\ar[r]\ar[d] & \tilde P_{m+1}^I\ar[r]\ar[d] & 0\\
  & & 0& 0&}\]
  The image of the inclusion $\tilde P_{m-1}\into \tilde P_{m,j}$ in the bottom row is the subrepresentation $\tilde P_{m-1,(j,j)}\subset\tilde P_{m,j}$ and part (1) follows.

  The first claim of part (2) is then immediate by repeatedly applying part (1) while the final claim of part (2) is a consequence of the definition of $\delta$.
\end{proof}

@@

Thus it remains to show that this cell decomposition of $\Gr^{\tilde Q}_\tbfe(\tilde P_{m}^{\uj'})$ is compatible with $\pi_{m,\uj'}^{\bf j}$ where ${\bf j}=(\underline{j_1},\underline{j_2},\ldots,\underline{j_r})$ with $\underline{j_1}=(\uj',j_{1,l},\ldots,j_{1,n-1})\in I_{n-1}$ and $j_{1,l}=j_l$. Recall that we have $\pi_{m,\uj'}^{\bf j}=\pi_{m,\uj}^{\bf j}\circ\pi_{m,\uj'}^{j_l}$ by Lemma \ref{le:truncated quotients}.(2).
